PANCHANGE_data_preprocessing

### 2020-08-04 PANCHANGE_data_preprocessing2020-07-15.Rmd
read in respiratory data and created chunks to decide on cut-offs for algoirthm symptoms

#Packages
Install packages (if they are not available in your version of R)
```{r Installing packages}
#install.packages("summarytools")
#install.packages("tidyverse")
#install.packages("psych")
```

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(tidyverse)
library(psych)
```

@@

@@

# glad ocir

```{r glad ocir cols}
glad.ocir.raw <- readRDS(file = paste0(data.raw_path, "/2020-08-03/glad/ocir_glad.rds"))
dim(glad.ocir.raw)
colnames(glad.ocir.raw)
```

```{r glad ocir cleaning}
glad.ocir.raw.id <- glad.ocir.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.glad = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    ocir.i_repeatedly_check_doors_windows_drawers_etc_prepan = ocir.i_repeatedly_check_doors_windows_drawers_etc.,
    ocir.arranged_things_change_upset_prepan = ocir.arranged_things_change_upset,
    ocir.repeat_feel_numbers_prepan = ocir.repeat_feel_numbers,
    ocir.clean_simply_wash_feel_prepan = ocir.clean_simply_wash_feel,
    ocir.mind_upset_unpleasant_thoughts_prepan = ocir.mind_upset_unpleasant_thoughts,
    ocir.afraid_avoid_throwing_things_prepan = ocir.afraid_avoid_throwing_things,
    ocir.light_switches_water_taps_prepan = ocir.light_switches_water_taps,
    ocir.arranged_things_prepan = ocir.arranged_things,
    ocir.good_feel_bad_numbers_prepan = ocir.good_feel_bad_numbers,
    ocir.hands_longer_wash_prepan = ocir.hands_longer_wash,
    ocir.saved_things_prepan = ocir.saved_things,
    ocir.frequently_difficulty_rid_nasty_prepan = ocir.frequently_difficulty_rid_nasty,
    ocir.i_check_things_more_often_than_necessary_prepan = ocir.i_check_things_more_often_than_necessary.,
    ocir.arranged_properly_objects_upset_prepan = ocir.arranged_properly_objects_upset,
    ocir.feel_compelled_count_things_prepan = ocir.feel_compelled_count_things,
    ocir.strangers_touch_touched_difficult_prepan = ocir.strangers_touch_touched_difficult,
    ocir.control_difficult_find_thoughts_prepan = ocir.control_difficult_find_thoughts,
    ocir.i_collect_things_i_dont_need_prepan = ocir.i_collect_things_i_dont_need
  ) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.ocir.raw.id)
# Inspect colnames
colnames(glad.ocir.raw.id)
#Differences
dim(glad.ocir.raw)[1]-dim(glad.ocir.raw.id)[1]
```

Calculate sum score for OCI-R

*** MD - need to adapt for participants with missing data (should be NA)
```{r ocir sum score prepan}
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
ocir.items_key_prepan <- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)

# Create vector with items that are included in sum score
ocir.items_prepan = c(
  "ocir.i_repeatedly_check_doors_windows_drawers_etc_prepan",
  "ocir.arranged_things_change_upset_prepan",
  "ocir.repeat_feel_numbers_prepan",
  "ocir.clean_simply_wash_feel_prepan",
  "ocir.mind_upset_unpleasant_thoughts_prepan",
  "ocir.afraid_avoid_throwing_things_prepan",
  "ocir.light_switches_water_taps_prepan",
  "ocir.arranged_things_prepan",
  "ocir.good_feel_bad_numbers_prepan",
  "ocir.hands_longer_wash_prepan",
  "ocir.saved_things_prepan",
  "ocir.frequently_difficulty_rid_nasty_prepan",
  "ocir.i_check_things_more_often_than_necessary_prepan",
  "ocir.arranged_properly_objects_upset_prepan",
  "ocir.feel_compelled_count_things_prepan",
  "ocir.strangers_touch_touched_difficult_prepan",
  "ocir.control_difficult_find_thoughts_prepan",
  "ocir.i_collect_things_i_dont_need_prepan"
  )

# Calulate sum score
ocir.scored_items_prepan <- scoreItems(keys = ocir.items_key_prepan,
                                        items = glad.ocir.raw.id[ocir.items_prepan],
                                        totals = TRUE, 
                                        missing = TRUE, 
                                        impute = 'none', 
                                        min = 0, 
                                        max = 4)
# Add in column to tibble
glad.ocir.raw.id$ocir.sum_score_prepan <- ocir.scored_items_prepan$scores

# Look at values
glad.ocir.raw.id %>%
  descr(ocir.sum_score_prepan,
        stats = "common"
        )
```

```{r glad ocir head}
head(glad.ocir.raw.id)
```

```{r glad ocir summary}
summarytools::dfSummary(
  glad.ocir.raw.id,
  graph.col = F,
  valid.col = F
  )
```

# COPING GLAD: GAD-7

@@

# Create vector with items that are included in total score
phq.items_retro = c(
  "phq.little_interest_or_pleasure_in_doing_things_retro_numeric",
"phq.feeling_down_depressed_or_hopeless_retro_numeric",
"phq.staying_asleep_sleeping_trouble_retro_numeric",
"phq.feeling_tired_or_having_little_energy_retro_numeric",
"phq.poor_appetite_or_overeating_retro_numeric",
"phq.feeling_bad_failure_family_retro_numeric",
"phq.trouble_concentrating_reading_newspaper_retro_numeric",
"phq.moving_fidgety_noticed_opposite_retro_numeric",
"phq.dead_hurting_thoughts_retro_numeric"
)

@@

# Create vector with items that are included in total score
phq.items_base = c(
  "phq.little_interest_or_pleasure_in_doing_things_base_numeric",
"phq.feeling_down_depressed_or_hopeless_base_numeric",
"phq.staying_asleep_sleeping_trouble_base_numeric",
"phq.feeling_tired_or_having_little_energy_base_numeric",
"phq.poor_appetite_or_overeating_base_numeric",
"phq.feeling_bad_failure_family_base_numeric",
"phq.trouble_concentrating_reading_newspaper_base_numeric",
"phq.moving_fidgety_noticed_opposite_base_numeric",
"phq.dead_hurting_thoughts_base_numeric"
)

@@

# glad COPING ocir

```{r coping glad ocir cols}
coping.glad.ocir.raw <- readRDS(file = paste0(data.raw_path, "/2020-08-03/glad/ocir_coping_glad.rds"))
dim(coping.glad.ocir.raw)
colnames(coping.glad.ocir.raw)
```

```{r coping glad ocir summary labels}
summarytools::dfSummary(
  coping.glad.ocir.raw,
  graph.col = F,
  valid.col = F
  #labels.col = F
)
```
```{r coping glad ocir cleaning}
coping.glad.ocir.raw.id <- coping.glad.ocir.raw %>%
    ocir.i_repeatedly_check_doors_windows_drawers_etc_retro = ocir.i_repeatedly_check_doors_windows_drawers_etc.,
    ocir.arranged_things_change_upset_retro = ocir.arranged_things_change_upset,
    ocir.repeat_feel_numbers_retro = ocir.numbers_repeat_feel, ## Different from GLAD prepandemic
    ocir.clean_simply_wash_feel_retro = ocir.feel_contaminated_clean_simply, ## Different from GLAD prepandemic
    ocir.mind_upset_unpleasant_thoughts_retro = ocir.mind_upset_unpleasant_thoughts,
    ocir.afraid_avoid_throwing_things_retro = ocir.afraid_avoid_throwing_things,
    ocir.light_switches_water_taps_retro = ocir.light_switches_water_taps,
    ocir.arranged_things_retro = ocir.arranged_things,
    ocir.good_feel_bad_numbers_retro = ocir.bad_numbers_good_feel, ## Different from GLAD prepandemic
    ocir.hands_longer_wash_retro = ocir.longer_hands_wash, ## Different from GLAD prepandemic
    ocir.saved_things_retro = ocir.saved_things,
    ocir.frequently_difficulty_rid_nasty_retro = ocir.frequently_rid_difficulty_nasty, ## Different from GLAD prepandemic
    ocir.i_check_things_more_often_than_necessary_retro = ocir.i_check_things_more_often_than_necessary.,
    ocir.arranged_properly_objects_upset_retro = ocir.arranged_properly_objects_upset,
    ocir.feel_compelled_count_things_retro = ocir.feel_compelled_count_things,
    ocir.strangers_touch_touched_difficult_retro = ocir.strangers_touch_touched_object, ## Different from GLAD prepandemic
    ocir.control_difficult_find_thoughts_retro = ocir.control_find_difficult_thoughts, ## Different from GLAD prepandemic
    ocir.i_collect_things_i_dont_need_retro = ocir.i_collect_things_i_dont_need., ## Different from GLAD prepandemic
    ocir.i_repeatedly_check_doors_windows_drawers_etc_base = ocir.i_repeatedly_check_doors_windows_drawers_etc..1,
    ocir.arranged_things_change_upset_base = ocir.arranged_things_change_upset.1,
    ocir.repeat_feel_numbers_base = ocir.numbers_repeat_feel.1, ## Different from GLAD prepandemic
    ocir.clean_simply_wash_feel_base = ocir.feel_contaminated_clean_simply.1, ## Different from GLAD prepandemic
    ocir.mind_upset_unpleasant_thoughts_base = ocir.mind_upset_unpleasant_thoughts.1,
    ocir.afraid_avoid_throwing_things_base = ocir.afraid_avoid_throwing_things.1,
    ocir.light_switches_water_taps_base = ocir.light_switches_water_taps.1,
    ocir.arranged_things_base = ocir.arranged_things.1,
    ocir.good_feel_bad_numbers_base = ocir.bad_numbers_good_feel.1, ## Different from GLAD prepandemic
    ocir.hands_longer_wash_base = ocir.longer_hands_wash.1, ## Different from GLAD prepandemic
    ocir.saved_things_base = ocir.saved_things.1,
    ocir.frequently_difficulty_rid_nasty_base = ocir.frequently_rid_difficulty_nasty.1, ## Different from GLAD prepandemic
    ocir.i_check_things_more_often_than_necessary_base = ocir.i_check_things_more_often_than_necessary..1,
    ocir.arranged_properly_objects_upset_base = ocir.arranged_properly_objects_upset.1,
    ocir.feel_compelled_count_things_base = ocir.feel_compelled_count_things.1,
    ocir.strangers_touch_touched_difficult_base = ocir.strangers_touch_touched_object.1, ## Different from GLAD prepandemic
    ocir.control_difficult_find_thoughts_base = ocir.control_find_difficult_thoughts.1, ## Different from GLAD prepandemic
    ocir.i_collect_things_i_dont_need_base = ocir.i_collect_things_i_dont_need..1

# Inspect dimensions
dim(coping.glad.ocir.raw.id)
# Inspect colnames
colnames(coping.glad.ocir.raw.id)
#Differences
dim(coping.glad.ocir.raw)[1]-dim(coping.glad.ocir.raw.id)[1]
```

Calculate sum score for OCI-R retro
```{r coping ocir sum score retro}
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
ocir.items_key_retro <- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)

# Create vector with items that are included in total score
ocir.items_retro = c(
  "ocir.i_repeatedly_check_doors_windows_drawers_etc_retro",
  "ocir.arranged_things_change_upset_retro",
  "ocir.repeat_feel_numbers_retro",
  "ocir.clean_simply_wash_feel_retro",
  "ocir.mind_upset_unpleasant_thoughts_retro",
  "ocir.afraid_avoid_throwing_things_retro",
  "ocir.light_switches_water_taps_retro",
  "ocir.arranged_things_retro",
  "ocir.good_feel_bad_numbers_retro",
  "ocir.hands_longer_wash_retro",
  "ocir.saved_things_retro",
  "ocir.frequently_difficulty_rid_nasty_retro",
  "ocir.i_check_things_more_often_than_necessary_retro",
  "ocir.arranged_properly_objects_upset_retro",
  "ocir.feel_compelled_count_things_retro",
  "ocir.strangers_touch_touched_difficult_retro",
  "ocir.control_difficult_find_thoughts_retro",
  "ocir.i_collect_things_i_dont_need_retro"
)

# Calulate total score
ocir.scored_items_retro <- scoreItems(keys = ocir.items_key_retro,
                                     items = coping.glad.ocir.raw.id[ocir.items_retro],
                                     totals = TRUE, 
                                     missing = TRUE, 
                                     impute = 'none', 
                                     min = 0, 
                                     max = 4)
# Add in column to tibble
coping.glad.ocir.raw.id$ocir.sum_score_retro <- ocir.scored_items_retro$scores

# Look at values
coping.glad.ocir.raw.id %>%
  descr(ocir.sum_score_retro,
        stats = "common"
  )
```

Calculate sum score for OCI-R base
```{r coping ocir sum score base}
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
ocir.items_key_base <- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)

# Create vector with items that are included in total score
ocir.items_base = c(
  "ocir.i_repeatedly_check_doors_windows_drawers_etc_base",
  "ocir.arranged_things_change_upset_base",
  "ocir.repeat_feel_numbers_base",
  "ocir.clean_simply_wash_feel_base",
  "ocir.mind_upset_unpleasant_thoughts_base",
  "ocir.afraid_avoid_throwing_things_base",
  "ocir.light_switches_water_taps_base",
  "ocir.arranged_things_base",
  "ocir.good_feel_bad_numbers_base",
  "ocir.hands_longer_wash_base",
  "ocir.saved_things_base",
  "ocir.frequently_difficulty_rid_nasty_base",
  "ocir.i_check_things_more_often_than_necessary_base",
  "ocir.arranged_properly_objects_upset_base",
  "ocir.feel_compelled_count_things_base",
  "ocir.strangers_touch_touched_difficult_base",
  "ocir.control_difficult_find_thoughts_base",
  "ocir.i_collect_things_i_dont_need_base"
)

# Calulate total score
ocir.scored_items_base <- scoreItems(keys = ocir.items_key_base,
                                    items = coping.glad.ocir.raw.id[ocir.items_base],
                                    totals = TRUE, 
                                    missing = TRUE, 
                                    impute = 'none', 
                                    min = 0, 
                                    max = 4)
# Add in column to tibble
coping.glad.ocir.raw.id$ocir.sum_score_base <- ocir.scored_items_base$scores

# Look at values
coping.glad.ocir.raw.id %>%
  descr(ocir.sum_score_base,
        stats = "common"
  )
```

```{r coping glad ocir head}
head(coping.glad.ocir.raw.id)
```

```{r coping glad ocir summary}
summarytools::dfSummary(
  coping.glad.ocir.raw.id,
  graph.col = F,
  valid.col = F,
  labels.col = F)
```

@@

glad data frame list
```{r glad data frame list}
glad.dfs.list <- list(
  glad.dem.raw.id,
  glad.gad.raw.id,
  glad.mhq.raw.categories,
  glad.ocir.raw.id,
  glad.pcl.raw.id,
  glad.phq.raw.id,
  coping.glad.gad.raw.id,
  coping.glad.pcl.raw.id,
  coping.glad.phq.raw.id,
  coping.glad.ocir.raw.id
)
```

@@

##Missingness of questionnaire items
#PHQ - baseline
PHQ Missingness NAs per person - count and percentages
```{r PHQ baseline missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_phq_base =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the PHQ items at baseline
           %in%
             phq.items_base]
        )
    )
  )

@@

#GAD - baseline
GAD Missingness NAs per person - count and percentages
```{r gad baseline missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_gad_base =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the gad items at baseline
           %in%
             gad.items_base]
        )
    )
  )

@@

#PCL - baseline
PCL Missingness NAs per person - count and percentages
```{r pcl baseline missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_pcl_base =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the pcl items at baseline
           %in%
             pcl.items_base]
        )
    )
  )

@@

#OCI-R - baseline
OCI-R Missingness NAs per person - count and percentages
```{r OCI-R baseline missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_ocir_base =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the OCI-R items at baseline
           %in%
             ocir.items_base]
        )
    )
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_ocir_base) #have a look at the frequency

dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>% #NA per person as a percentage
  mutate(
    miss_pc_total_ocir_base = 
      na_per_person_ocir_base/18 # NA per person divided by the amount of items
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$miss_pc_total_ocir_base)
```

#OCI-R - retrospective
OCI-R Missingness NAs per person - count and percentages
```{r OCI-R retrospective missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_ocir_retro =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the OCI-R items at retroline
           %in%
             ocir.items_retro]
        )
    )
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_ocir_retro) #have a look at the frequency

dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>% #NA per person as a percentage
  mutate(
    miss_pc_total_ocir_retro = 
      na_per_person_ocir_retro/18 # NA per person divided by the amount of items
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$miss_pc_total_ocir_retro)
```

#OCI-R - prepandemic
OCI-R Missingness NAs per person - count and percentages
```{r OCI-R prepan missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_ocir_prepan =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the OCI-R items at prepanline
           %in%
             ocir.items_prepan]
        )
    )
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_ocir_prepan) #have a look at the frequency

dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>% #NA per person as a percentage
  mutate(
    miss_pc_total_ocir_prepan = 
      na_per_person_ocir_prepan/18 # NA per person divided by the amount of items
  )

freq(dat.no.sex.no.gender.no.age.no.ibd$miss_pc_total_ocir_prepan)
```

Count for OCI-R baseline, retrospective, and prepandemic missingness per person
```{r count of people missing a number of items for OCI-R}
#baseline
table(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_ocir_base)
#retrospective
table(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_ocir_retro)
#pre-pandemic
table(dat.no.sex.no.gender.no.age.no.ibd$na_per_person_ocir_prepan)
```

#Creating "complete cases" datasets  - NOTE THIS PROBABLY WONT BE USED BUT JUST TO GIVE AN IDEA OF COMPLETE CASES NUMBERS. KT. 
Complete case data for PHQ, GAD and PCL - 3345 remaining, drop 35,036.
```{r complete cases only}
all.items <- c(phq.items_prepan,
                phq.items_base,
                phq.items_retro,
                gad.items_prepan,
                gad.items_base,
                gad.items_retro,
                pcl.items_prepan,
                pcl.items_base,
                pcl.items_retro)

dat.no.sex.no.gender.no.age.no.ibd.complete.cases <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, all.items ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases)[1]
```

PHQ
Complete case data for PHQ baseline (COPING) - 9,260 remaining
```{r complete cases phq base}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.base <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, phq.items_base ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.base)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.base)[1]
```

Complete case data for PHQ pre-pandemic (GLAD) - 37,308 remaining
```{r complete cases phq prepan}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.prepan <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, phq.items_prepan ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.prepan)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.prepan)[1]
```

Complete case data for PHQ retrospective (COPING) - 9,262 remaining
```{r complete cases phq retro}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.retro <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, phq.items_retro ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.retro)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.retro)[1]
```

GAD
Complete case data for GAD baseline (COPING) - 12,457 remaining
```{r complete cases gad base}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.base <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, gad.items_base ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.base)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.base)[1]
```

Complete case data for gad pre-pandemic (GLAD) - 36,149 remaining
```{r complete cases gad prepan}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.prepan <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, gad.items_prepan ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.prepan)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.prepan)[1]
```

Complete case data for gad retrospective (COPING) - 8,930 remaining
```{r complete cases gad retro}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.retro <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, gad.items_retro ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.retro)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.retro)[1]
```

>>>>

#Packages
Install packages (if they are not available in your version of R)
```{r Installing packages}
#install.packages("summarytools")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("dplyr")
```

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(dplyr)
library(tidyverse)
```

@@

#Add_numeric function
Function used to convert character variables into numeric variables. 
```{r add_numeric function}
add_numeric <- function(dat, exclude = NULL) {
  dat_fct <- sjlabelled::as_label(dat)

  dat <- dat[!colnames(dat) %in% exclude]
  colnames(dat) <- paste(colnames(dat), "numeric", sep = "_")
  return(bind_cols(dat_fct, dat))
}
```

@@

## COPING GLAD: GAD-7

@@

# Create vector with items that are included in total score
phq.items_retro <- c(
  "phq.little_interest_or_pleasure_in_doing_things_retro_numeric",
"phq.feeling_down_depressed_or_hopeless_retro_numeric",
"phq.staying_asleep_sleeping_trouble_retro_numeric",
"phq.feeling_tired_or_having_little_energy_retro_numeric",
"phq.poor_appetite_or_overeating_retro_numeric",
"phq.feeling_bad_failure_family_retro_numeric",
"phq.trouble_concentrating_reading_newspaper_retro_numeric",
"phq.moving_fidgety_noticed_opposite_retro_numeric",
"phq.dead_hurting_thoughts_retro_numeric"
)

@@

# Create vector with items that are included in total score
phq.items_base <- c(
  "phq.little_interest_or_pleasure_in_doing_things_base_numeric",
"phq.feeling_down_depressed_or_hopeless_base_numeric",
"phq.staying_asleep_sleeping_trouble_base_numeric",
"phq.feeling_tired_or_having_little_energy_base_numeric",
"phq.poor_appetite_or_overeating_base_numeric",
"phq.feeling_bad_failure_family_base_numeric",
"phq.trouble_concentrating_reading_newspaper_base_numeric",
"phq.moving_fidgety_noticed_opposite_base_numeric",
"phq.dead_hurting_thoughts_base_numeric"
)

@@

@@

glad data frame list
```{r glad data frame list}
glad.dfs.list <- list(
  glad.dem.raw.id,
  glad.gad.raw.id,
  glad.mhq.raw.categories,
#  glad.ocir.raw.id,
  glad.pcl.raw.id,
  glad.phq.raw.id,
  coping.glad.gad.raw.id,
  coping.glad.pcl.raw.id,
  coping.glad.phq.raw.id
)
```

@@

##Missigness of questionnaire items
#PHQ - baseline
PHQ Missingness NAs per person - count and percentages
```{r PHQ basline missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_phq_base =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the PHQ items at baseline
           %in%
             phq.items_base]
        )
    )
  )

@@

#GAD - baseline
GAD Missingness NAs per person - count and percentages
```{r gad basline missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_gad_base =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the gad items at baseline
           %in%
             gad.items_base]
        )
    )
  )

@@

#PCL - baseline
PCL Missingness NAs per person - count and percentages
```{r pcl basline missingness}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    na_per_person_pcl_base =         # variable for NA per person
      rowSums(                       
        is.na(dat.no.sex.no.gender.no.age.no.ibd # sum the rows that are NA
          [,colnames(dat.no.sex.no.gender.no.age.no.ibd) #for all the rows and columns that are in the pcl items at baseline
           %in%
             pcl.items_base]
        )
    )
  )

@@

#Sum scores missingness cut offs
KT: This series of chunks creates a "full" sum score for the PHQ, GAD and PCL.

PHQ
The PHQ sum score is set to NA if they have more than one missing item, and it is not the last suicide ideation item (they do not meet the cut off criteria - phq.base_cut_off)
```{r sum score cut offs PHQ base}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    phq.sum_score_base_full =
      if_else(na_per_person_phq_base > 1 & phq.base_cut_off == 0,
              NA_real_,
              phq.sum_score_base
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$phq.sum_score_base)
summary(dat.no.sex.no.gender.no.age.no.ibd$phq.sum_score_base_full)
```

```{r sum score cut offs PHQ prepan}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    phq.sum_score_prepan_full =
      if_else(na_per_person_phq_prepan > 1 & phq.prepan_cut_off == 0,
              NA_real_,
              phq.sum_score_prepan
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$phq.sum_score_prepan)
summary(dat.no.sex.no.gender.no.age.no.ibd$phq.sum_score_prepan_full)
```

```{r sum score cut offs PHQ retro}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    phq.sum_score_retro_full =
      if_else(na_per_person_phq_retro > 1 & phq.retro_cut_off == 0,
              NA_real_,
              phq.sum_score_retro
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$phq.sum_score_retro)
summary(dat.no.sex.no.gender.no.age.no.ibd$phq.sum_score_retro_full)
```

GAD
The GAD sum score is set to NA if they have any missing items
```{r sum score cut offs GAD base}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    gad.sum_score_base_full =
      if_else(na_per_person_gad_base > 0,
              NA_real_,
              gad.sum_score_base
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$gad.sum_score_base)
summary(dat.no.sex.no.gender.no.age.no.ibd$gad.sum_score_base_full)
```

```{r sum score cut offs GAD prepan}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    gad.sum_score_prepan_full =
      if_else(na_per_person_gad_prepan > 0,
              NA_real_,
              gad.sum_score_prepan
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$gad.sum_score_prepan)
summary(dat.no.sex.no.gender.no.age.no.ibd$gad.sum_score_prepan_full)
```

```{r sum score cut offs GAD retro}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    gad.sum_score_retro_full =
      if_else(na_per_person_gad_retro > 0,
              NA_real_,
              gad.sum_score_retro
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$gad.sum_score_retro)
summary(dat.no.sex.no.gender.no.age.no.ibd$gad.sum_score_retro_full)
```

PCL
The PCL sum score is set to NA if they have any missing items - only for basline and prepan.
```{r sum score cut offs PCL base}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    pcl.sum_score_base_full =
      if_else(na_per_person_pcl_base > 0,
              NA_real_,
              pcl.sum_score_base
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$pcl.sum_score_base)
summary(dat.no.sex.no.gender.no.age.no.ibd$pcl.sum_score_base_full)
```

```{r sum score cut offs PCL prepan}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    pcl.sum_score_prepan_full =
      if_else(na_per_person_pcl_prepan > 0,
              NA_real_,
              pcl.sum_score_prepan
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$pcl.sum_score_prepan)
summary(dat.no.sex.no.gender.no.age.no.ibd$pcl.sum_score_prepan_full)
```
